---
title: "Homework 3"
author: "Trevor Caughlin"
date: "January 29, 2018"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1: What is the effect of cut quality on the price of 50,000 round cut diamonds?
These data are the prices of 50,000 round cut diamonds (source: http://ggplot2.tidyverse.org/reference/diamonds.html )

Data includes:
**Price** in US dollars (\$326--\$18,823)
**Carat** (weight) of the diamond (0.2--5.01)
**Cut** (quality of the cut; Fair, Good, Very Good, Premium, Ideal)


```{r diamonds}
diamonds<-read.csv("diamond.csv")
head(diamonds)
```

What is the effect of each cut on the price of a typical diamond (in dollars)?

As output, produce a boxplot of results, parameter estimates, and 95% CI

#Question 2: Does education have an impact on contraception use?

These data represent current contraception use of in a sample of 1607 married women from Fiji. Source: "Reference: Little, R. J. A. (1978). Generalized Linear Models for Cross-Classified Data from the WFS. World Fertility Survey Technical Bulletins, Number 5."

Data inclues: 
**age** cohort of women
**education** education level
**notUsing** count of women in each sample not using contraception
**using** count of women in each sample using contraception
**total number of women in sample**

```{r cuse}
cuse<-read.csv("contraception.csv")
head(cuse)
```

Use a binomial glm to test the hypothesis that increased education promotes contraception use.

As output, produce a boxplot of results, parameter estimates, and 95% CI (and a short sentence on whether you reject the null hypothesis or not).


#Question 3
A controversial recent paper claimed that female-named hurricanes led to more deaths than male-named hurricanes (also known as himmicanes), because people do not take hurricanes as seriously as himmicanes. You can find a link to the paper here: http://www.pnas.org/content/111/24/8782.short

Citation: Jung, Kiju, Sharon Shavitt, Madhu Viswanathan, and Joseph M. Hilbe. 2014. "Female Hurricanes Are Deadlier than Male Hurricanes." Proceedings of the National Academy of Sciences 111 (24): 8782-87. doi:10.1073/pnas.1402786111.

Here are the data used for their paper:

```{r hurricanes}
hurricanes<-read.csv("Hurricane Dataset.csv")
head(hurricanes)
```

Re-analyze their data using Poisson glm, with a focus on **alldeaths** as a response variable and **Gender_MF** as a predictor variable. How do your results differ from Jung et al.'s conclusions? What could Jung et al. have done differently to inspire more confidence in their analyses?

#Question 4
Find one dataset from your own research that could be modelled as normal, binomial, or poisson regression (note: http://datadryad.org/ is a great source for existing datasets if you don't have your own data yet). Run a glm on your data and interpret results in terms of effect size, visualization, and statistical significance.